---
title: "PHW251 Problem Set 4"
author: "Clara Voong"
date: "9/29/2024"
output: pdf_document
---

For this problem set you will tidy up a dataset of 500 individuals. We also want to calculate each individual's BMI and appropriately categorize them.

Load your data (500_Person_Gender_Height_Weight.csv):

```{r load, include=FALSE}
library(tidyverse)
library(magrittr)

data <- read_csv("~/phw251_fall24_cv/week 5/500_Person_Gender_Height_Weight.csv")
```

### Question 1

Clean the column headers to be all lower case, have no spaces, and rename "Location information" to location.

```{r clean}
data <- data %>%
  rename(
    location = `Location information`,
    gender = GENDER
  )

colnames(data)
```

\newpage 

### Question 2

Create a new variable that calculates BMI for each individual.

You will need to navigate the different system of measurements (metric vs imperial). Only the United States is using imperial.

* BMI calculation and conversions:
    - metric:   $BMI = weight (kg) / [height (m)]^2$
    - imperial: $BMI = 703 * weight (lbs) / [height (in)]^2$
    - 1 foot = 12 inches
    - 1 cm = 0.01 meter

Although there's many ways you can accomplish this task, we want you to use an if_else() to calculate BMI with the appropriate formula based on each person's location. 

```{r bmi_calc}
# head(data)

data <- data %>%
  mutate(
    bmi = if_else(
      location %in% c("United Kingdom","Taiwan"), 
      weight/((height*0.01)^2), 
      703 * weight/((height*12)^2)
    )
  )
head(data)
```

\newpage

### Question 3

Create a new variable that categorizes BMI with case_when():

* Underweight: BMI below 18.5
* Normal: 18.5-24.9
* Overweight: 25.0-29.9
* Obese: 30.0 and Above

```{r bmi_cat}
data <- data %>% mutate(
  bmi_cat = 
    case_when(
      bmi < 18.5 ~ "Underweight",
      bmi <= 24.9 ~ "Normal",
      bmi <= 29.9 ~ "Overweight",
      TRUE ~ "Obese"
    )
)

head(data)
```

Could we have used if_else()?  

Yes, but we would have the specify the lower bound as well for each line in contrast with case_when(), 
where the lines are run sequentially and the first match in the function would be the output.

\newpage

### Question 4

Arrange your data first by location and then by descending order of BMI.

```{r arrange_ds}
data <- data %>%
  arrange(location, desc(bmi))

head(data)
```

### Question 5

Use a dplyr method to remove the height, weight, and BMI columns from your data.

```{r remove_cols}
# colnames(data)

data <- data %>% select(location, gender, bmi_cat)

head(data)
```

\newpage

### Optional Challenge

Perform all the actions in this problem set with one dpylr call. 

```{r mega_call}
data <- read_csv("~/phw251_fall24_cv/week 5/500_Person_Gender_Height_Weight.csv") %>%  
  rename(
    location = `Location information`,
    gender = GENDER
  ) %>%
  mutate(
    bmi = if_else(
      location %in% c("United Kingdom","Taiwan"), 
      weight/((height*0.01)^2), 
      703 * weight/((height*12)^2)
    )
  ) %>%
  mutate(
  bmi_cat = 
    case_when(
      bmi < 18.5 ~ "Underweight",
      bmi <= 24.9 ~ "Normal",
      bmi <= 29.9 ~ "Overweight",
      TRUE ~ "Obese"
    )
) %>%
  arrange(location, desc(bmi)) %>%
  select(location, gender, bmi_cat)
  
head(data)
```
